---
title: "wsm_modelS_testing"
format: html
---

```{r}
#| output: false 
source(here::here("wsm_model_zscore.R"))
source(here::here("wsm_model_MinMax.R"))
library(dplyr)
library(tidyverse)
library(sf)
library(ggplot2)
library(plotly)
library(tidyr)
```

## Fake dataset

Building a fake dataset that I can predict the outcomes

```{r}
dam_df <- data.frame(
  dam_id = c("2001", "2002", "2003", "2004", "2005", "2006", "2007", "2008", "2009", "2010"), 
  bb_id = c(1, 1, 1, 2, 2, 2, 3, 3, 4, 4), # trunk id 
  min_distance_upstream_km = c(NA, 10, 30, NA, 3, NA, 30, 200, NA, 2),  
  min_distance_downstream_km = c(20, 10, NA, NA, 15, 30, 100, 10, NA, 14), 
  # connectivity indexes 
  cascade_status = c(0,1,0,NA,1,0,0,0,NA,1), 
  csi = c(0.40, 0.10, 0.22, 0.90, 0.12, 0.25, 0.81, 0.70, 0.95, 0.07),
  # fatal flaws 
  in_protected_area = c(0,1,1,0,0,1,1,1,0,0) # CHANGED to have 1 = good (no PA) and 0 = bad (inside PA)... WAS c(1,0,0,1,1,0,0,0,1,1) 
)


# upsteam, downstream, cascade 
dam_df <- dam_df %>% 
  mutate(connectivity_category = case_when(
    min_distance_upstream_km <= 15 & min_distance_downstream_km <= 15 ~ "cascade", 
    is.na(min_distance_upstream_km) & is.na(min_distance_downstream_km) ~ "FFR",
    is.na(min_distance_upstream_km) & min_distance_downstream_km <= 50 | 
      min_distance_upstream_km > 50 & min_distance_downstream_km <= 50 ~ "upstream", 
    min_distance_upstream_km <= 50 & is.na(min_distance_downstream_km) | 
      min_distance_upstream_km <= 50 & min_distance_downstream_km > 50 ~ "downstream", 
    TRUE ~ NA
  ))


# Change the connectivity_category to numbers 
dam_df <- dam_df %>% 
  mutate(connectivity_category_num  = case_when(
    connectivity_category == "cascade" ~ 5, 
    connectivity_category == "upstream" ~ 3, 
    connectivity_category == "downstream" ~ 3, 
    connectivity_category == "ffr" ~ 1, 
    TRUE ~ 1
    ))


dam_df

```

Trying function on fake data

```{r}

criteria <- c("csi", "connectivity_category_num", "in_protected_area")

criteria_weights <- c(csi = 0.5, connectivity_category_num = 0.3, in_protected_area = 0.2)

criteria_type <- c(csi = "nonbeneficial", connectivity_category_num = "beneficial", in_protected_area = "beneficial")

# Min Max WSM 
wsm_MinMax_results <- wsm_model_MinMax(dataset = dam_df, 
     criteria = criteria, 
     criteria_weights = criteria_weights, 
     criteria_type = criteria_type)

# Z score WSM 
wsm_Zscore_results <-  wsm_model_zscore(dataset = dam_df, 
                                        criteria = criteria, 
                                        criteria_weights = criteria_weights, 
                                        criteria_type = criteria_type)
```

Show results with a boxplot

```{r}

wsm_MinMax_results %>% 
  ggplot() + 
  geom_boxplot(aes(x = wsm_scores, y = connectivity_category)) + 
  theme_gray() + 
  labs(title = "Boxplot of MinMax WSM results per catagory")

wsm_Zscore_results %>% 
  ggplot() + 
  geom_boxplot(aes(x = wsm_scores, y = connectivity_category)) + 
  theme_gray() + 
  labs(title = "Boxplot of Zscore WSM results per catagory")


```

The main difference is the X-axis ranges.

## Trying real Dataset

**Data used:** Dam connectivity of the Mekong Basin.

-   `conn_v3` is from Connectivity repo's `function_testing_v3.qmd`. Before running `function_testing_v3.qmd`, need to run net_with_dams_from_rivernetwork.

-   `protected_areas` from the protected Area repo.

```{r}
# Load data 
reach_df <- conn_v3$reach_df # From `function_testing_v3.qmd` 
protected_areas <- readRDS('../../../../../capstone/netzerohydro/data/cleaned_v1/protected_area_draft.rds')

# Add Protected areas 
protected_areas$dam_id <- as.character(protected_areas$dam_id) # Fix dam_id 
reach_df <- reach_df %>% separate_wider_delim(dam_id, "_", names = c("dam_dataset", "dam_id")) # Seperate dam_id column to match the pa 
reach_df_pa <- left_join(reach_df, protected_areas, by = 'dam_id') # Join PA with dam data 

# Add CSI 
# Join river-edge attributes by bb_id.
mekong_dams_final <- add_ffr_attr(
  out$net_with_dams,
  reach_df_pa,
  edge_attr_cols = c("csi", "ord_stra", "bas_name"),
  future_extra   = future_dams %>%
  dplyr::mutate(dam_id = as.character(.data$dam_id)) %>%
  dplyr::select(dam_id, capacity_mw, project_name)
)
```

```{r}
mekong_dams_final <- mekong_dams_final %>% 
  mutate(connectivity_category_num = case_when(
    connectivity_category == "cascade_classic" ~ 7, 
    connectivity_category == "cascade1" ~ 6,
    connectivity_category == "cascade2" ~ 5, 
    connectivity_category == "upstream" ~ 3,
    connectivity_category == "downstream" ~ 3,
    connectivity_category == "FFR" ~ 1,
  )) %>% 
  mutate(in_pa_num = ifelse(in_protected_area == FALSE, 1, 0))  # 1 = No in PA (good)
  #st_drop_geometry() 
```

Run models

```{r}

criteria <- c('csi', 'connectivity_category_num', 'in_pa_num')

criteria_weights <- c(csi = 0.4, connectivity_category_num = 0.5, in_pa_num = 0.1)

criteria_type <- c(csi = "nonbeneficial", # Lower number = better 
                       connectivity_category_num = "beneficial", # Higher number = better 
                       in_pa_num = "beneficial")


wsm_MinMax_results2 <- wsm_model_MinMax(dataset = mekong_dams_final, 
     criteria = criteria, 
     criteria_weights = criteria_weights, 
     criteria_type = criteria_type)

wsm_Zscore_results2 <- wsm_model_zscore(dataset = mekong_dams_final, 
     criteria = criteria, 
     criteria_weights = criteria_weights, 
     criteria_type = criteria_type)

```

Show results with a boxplot

```{r}

wsm_MinMax_results2 %>% 
  ggplot() + 
  geom_boxplot(aes(x = wsm_scores, y = connectivity_category)) + 
  theme_gray() + 
  labs(title = "Boxplot of MinMax WSM results per catagory")

wsm_Zscore_results2 %>% 
  ggplot() + 
  geom_boxplot(aes(x = wsm_scores, y = connectivity_category)) + 
  theme_gray() + 
  labs(title = "Boxplot of Zscore WSM results per catagory")

```

```{r}
# Factor the connectivity_category to get a specific order 
wsm_MinMax_results2$connectivity_category <- factor(wsm_MinMax_results2$connectivity_category, levels = c('cascade_classic', 'cascade1', 'cascade2', 'upstream', 'downstream', "FFR"))
wsm_Zscore_results2$connectivity_category <- factor(wsm_Zscore_results2$connectivity_category, levels = c('cascade_classic', 'cascade1', 'cascade2', 'upstream', 'downstream', "FFR"))

# Create a color paletes 
connectivity_cat_pal <- c("cascade_classic" = "#003800",
                          "cascade1" = "#1cbf3a",
                          "cascade2" = "#b7ffd7",
                          "upstream" = "#ffe500",
                          "downstream" = "#f99b2f",
                          "FFR" = "#c13000")


# Histogram 
wsm_MinMax_results2 %>% 
  #filter(connectivity_category == "cascade") %>% 
  ggplot() +
  geom_histogram(aes(wsm_scores, fill = connectivity_category)) + 
  theme_light() + 
  scale_fill_manual(values = connectivity_cat_pal) +
  labs(title = "MinMax WSM Scrores", 
       x = "WSM Scores", 
       y = "Count", 
       fill = "Connectivity Catagory") 


# Histogram 
wsm_Zscore_results2 %>% 
  #filter(connectivity_category == "cascade") %>% 
  ggplot() +
  geom_histogram(aes(wsm_scores, fill = connectivity_category)) + 
  theme_light() + 
  scale_fill_manual(values = connectivity_cat_pal) +
  labs(title = "Zscore WSM Scrores", 
       x = "WSM Scores", 
       y = "Count", 
       fill = "Connectivity Catagory") 

```

Look at results with a table wsm_Zscore_results2 does not have letters

```{r}
# Table
wsm_MinMax_results2 %>% 
  group_by(score_letter) %>% 
  summarise(count = n(), 
            avg_scores = mean(wsm_scores, na.rm = TRUE), 
            max_scores = max(wsm_scores, na.rm = TRUE), 
            min_scores = min(wsm_scores, na.rm = TRUE), 
            median_scores = median(wsm_scores, na.rm = TRUE), 
            avg_csi = mean(csi, na.rm = TRUE)) %>% 
  kableExtra::kbl(digits = 3) %>% 
  kableExtra::kable_styling(bootstrap_options = c("striped", "hover", "condensed"))


```

Look at results with a map

```{r}

# Make it sf and 
wsm_MinMax_results2_sf <- st_as_sf(wsm_MinMax_results2, crs = 4326)

# Check CRS 
st_crs(wsm_MinMax_results2_sf) == st_crs(east_as_basin_sf)
wsm_MinMax_results2_sf <- st_set_crs(wsm_MinMax_results2_sf, 4326) 


# Plot 
# GGplot 
wsm_MinMax_ggmap <- ggplot() +
  geom_sf(data = east_as_basin_sf, fill = NA, color = "black", alpha = 0.5) +
  geom_sf(data = wsm_MinMax_results2_sf, aes(color = wsm_scores, 
                                 text = paste0(
      "<b>Dam_ID: ", dam_id, "</b><br>", 
      "<b>CSI: </b>", csi, "<br>", 
      "<b>Connectivity Category: </b>", connectivity_category_num, "<br>", 
      "<b>In a PA: </b>", in_protected_area, "<br>",
      "<b>Capacity: </b>", capacity_mw, "mw"
    ))) +
  scale_color_gradientn(
    colors = colorRampPalette(c("#0f31bc", "#77d613", "orange"))(100),
    name = "WSM Scores"
  ) +
  #theme_minimal() + 
  theme_void() + 
  labs(title = "Weighted Sum Model Scores for Future Dams in Southeast Asia")

wsm_MinMax_ggmap_interactive <- ggplotly(wsm_MinMax_ggmap , 
         tooltip = "text" # inside of hover tool tip
         )

wsm_MinMax_ggmap_interactive
```

```{r}
# Make it sf and 
wsm_Zscore_results2_sf <- st_as_sf(wsm_Zscore_results2, crs = 4326)

# Check CRS 
st_crs(wsm_Zscore_results2_sf) == st_crs(east_as_basin_sf)
wsm_Zscore_results2_sf <- st_set_crs(wsm_Zscore_results2_sf, 4326) 


# Plot 
# GGplot 
wsm_Zscore_ggmap <- ggplot() +
  geom_sf(data = east_as_basin_sf, fill = NA, color = "black", alpha = 0.5) +
  geom_sf(data = wsm_Zscore_results2_sf, aes(color = wsm_scores, 
                                 text = paste0(
      "<b>Dam_ID: ", dam_id, "</b><br>", 
      "<b>CSI: </b>", csi, "<br>", 
      "<b>Connectivity Category: </b>", connectivity_category_num, "<br>", 
      "<b>In a PA: </b>", in_protected_area, "<br>",
      "<b>Capacity: </b>", capacity_mw, "mw"
    ))) +
  scale_color_gradientn(
    colors = colorRampPalette(c("#0f31bc", "#77d613", "orange"))(100),
    name = "WSM Scores"
  ) +
  #theme_minimal() + 
  theme_void() + 
  labs(title = "Weighted Sum Model Scores for Future Dams in Southeast Asia")

wsm_Zscore_ggmap_interactive <- ggplotly(wsm_Zscore_ggmap , 
         tooltip = "text" # inside of hover tool tip
         )

wsm_Zscore_ggmap_interactive
```
