---
title: "wsm_modelS_testing"
format: html
---

This quarto document runs the MDCA functions over a fake dataset and the hydropower dataset produced by the analysis with some additional data visualizations.  

Load Libraries
```{r}
#| output: false 
# The functions 
#source(here::here('R', "wsm_model_zscore.R"))
#source(here::here('R', "wsm_model_MinMax.R"))

# The packages 
library(dplyr)
library(tidyverse)
library(sf)
library(ggplot2)
library(plotly)
library(tidyr)
library(leaflet)
library(leaflet.extras)
```

## Fake dataset

Building a fake dataset in which I can predict the outcomes.

```{r}
dam_df <- data.frame(
  dam_id = c("2001", "2002", "2003", "2004", "2005", "2006", "2007", "2008", "2009", "2010"), 
  bb_id = c(1, 1, 1, 2, 2, 2, 3, 3, 4, 4), # trunk id 
  min_distance_upstream_km = c(NA, 10, 30, NA, 3, NA, 30, 200, NA, 2),  
  min_distance_downstream_km = c(20, 10, NA, NA, 15, 30, 100, 10, NA, 14), 
  # connectivity indexes 
  cascade_status = c(0,1,0,NA,1,0,0,0,NA,1), 
  csi = c(0.40, 0.10, 0.22, 0.90, 0.12, 0.25, 0.81, 0.70, 0.95, 0.07),
  # fatal flaws 
  in_protected_area = c(0,1,1,0,0,1,1,1,0,0) # CHANGED to have 1 = good (no PA) and 0 = bad (inside PA)... WAS c(1,0,0,1,1,0,0,0,1,1) 
)

# Adding upsteam, downstream, cascade classifications 
dam_df <- dam_df %>% 
  mutate(connectivity_category = case_when(
    min_distance_upstream_km <= 15 & min_distance_downstream_km <= 15 ~ "cascade", 
    is.na(min_distance_upstream_km) & is.na(min_distance_downstream_km) ~ "FFR",
    is.na(min_distance_upstream_km) & min_distance_downstream_km <= 50 | 
      min_distance_upstream_km > 50 & min_distance_downstream_km <= 50 ~ "upstream", 
    min_distance_upstream_km <= 50 & is.na(min_distance_downstream_km) | 
      min_distance_upstream_km <= 50 & min_distance_downstream_km > 50 ~ "downstream", 
    TRUE ~ NA
  ))


# Change the connectivity_category to numbers 
dam_df <- dam_df %>% 
  mutate(connectivity_category_num  = case_when(
    connectivity_category == "cascade" ~ 5, 
    connectivity_category == "upstream" ~ 3, 
    connectivity_category == "downstream" ~ 3, 
    connectivity_category == "ffr" ~ 1, 
    TRUE ~ 1
    ))


```

Running the functions on fake data

```{r}

criteria <- c("csi", "connectivity_category_num", "in_protected_area")

criteria_weights <- c(csi = 0.5, connectivity_category_num = 0.3, in_protected_area = 0.2)

criteria_type <- c(csi = "nonbeneficial", connectivity_category_num = "beneficial", in_protected_area = "beneficial")

# Min Max WSM 
wsm_MinMax_results <- wsm_model_MinMax(dataset = dam_df, 
     criteria = criteria, 
     criteria_weights = criteria_weights, 
     criteria_type = criteria_type)

# Z score WSM 
wsm_Zscore_results <-  wsm_model_zscore(dataset = dam_df, 
                                        criteria = criteria, 
                                        criteria_weights = criteria_weights, 
                                        criteria_type = criteria_type)

```

Show results of MCDA Min-Max and Z-scores with a boxplot - score distribution per connectivity category.

```{r}
# Box plots 
wsm_MinMax_results %>% 
  ggplot() + 
  geom_boxplot(aes(x = wsm_scores, y = connectivity_category)) + 
  theme_gray() + 
  labs(title = "Boxplot of MinMax WSM results per catagory")

wsm_Zscore_results %>% 
  ggplot() + 
  geom_boxplot(aes(x = wsm_scores, y = connectivity_category)) + 
  theme_gray() + 
  labs(title = "Boxplot of Zscore WSM results per catagory")


```

Scores are between models are very similar. Consistently, FFR scores the worst and Casacades score the best in the fake dataset. Min-Max normalization ranges the score from [0,1]. The MCDA Z-score range beyond [0,1]. 

## Trying With Real Dataset 

**Load data**

This data is from (1) running the connectivity repository workflow with GDW and FRHeD datasets, (2) getting the Protected area dataset from the fatal-flaw repository (3) Merging the two in the `prep_data.qmd` in this repository. 

```{r}
# out_enriched_pa from the `prep_data.qmd` repo 
out_enriched <- out_enriched_pa %>% filter(dam_type == "future")

# out_enriched <- readRDS(here::here("data", "out_enriched_v1.rds")) %>% filter(dam_type == "future") # If saved it as suggested in the `prep_data.qmd`
```

Run the model using set weights 
```{r}
criteria <- c('csi', 'connectivity_category_num', 'n_protected_areas')

# Can choose other weights if desired 
criteria_weights <- c(csi = 0.4, connectivity_category_num = 0.5, n_protected_areas = 0.1)

criteria_type <- c(csi = "nonbeneficial", # Lower number = better 
                       connectivity_category_num = "beneficial", # Higher number = better 
                       n_protected_areas = "beneficial")

# Run functions 
wsm_MinMax_results2 <- wsm_model_MinMax(dataset = out_enriched, 
     criteria = criteria, 
     criteria_weights = criteria_weights, 
     criteria_type = criteria_type)

wsm_Zscore_results2 <- wsm_model_zscore(dataset = out_enriched, 
     criteria = criteria, 
     criteria_weights = criteria_weights, 
     criteria_type = criteria_type)

```


Make catagory a factor and add color palette 
```{r}
# Factor the connectivity_category to get a specific order 
wsm_MinMax_results2$connectivity_category <- factor(wsm_MinMax_results2$connectivity_category, levels = c('cascade_classic', 'cascade1', 'cascade2', 'has_downstream_dam', 'has_upstream_dam', "undammed"))
wsm_Zscore_results2$connectivity_category <- factor(wsm_Zscore_results2$connectivity_category, levels = c('cascade_classic', 'cascade1', 'cascade2', 'has_downstream_dam', 'has_upstream_dam', "undammed"))

# Create a color palettes 
connectivity_cat_pal <- c("cascade_classic" = "#0e5faf",
                          "cascade1" = "#67A9CF",
                          "cascade2+" = "#D1E5F0",
                          "has_downstream_dam" = "#f7d35d",
                          "has_upstream_dam" = "#EF8A62", 
                          "undammed" = "#B2182B")


```

### Look at results with... 

A) Boxplots
```{r}
# Boxplots 
wsm_MinMax_results2 %>% 
  ggplot() + 
  geom_boxplot(aes(x = wsm_scores, y = connectivity_category)) + 
  theme_gray() + 
  labs(title = "Boxplot of MinMax WSM results per catagory")

wsm_Zscore_results2 %>% 
  ggplot() + 
  geom_boxplot(aes(x = wsm_scores, y = connectivity_category)) + 
  theme_gray() + 
  labs(title = "Boxplot of Zscore WSM results per catagory")

```


B) Histogram 
```{r}
# Histogram for MCDA Min Max results 
wsm_MinMax_results2 %>% 
  #filter(connectivity_category == "cascade") %>% 
  ggplot() +
  geom_histogram(aes(wsm_scores, fill = connectivity_category)) + 
  theme_light() + 
  scale_fill_manual(values = connectivity_cat_pal) +
  labs(title = "MinMax WSM Scores", 
       x = "WSM Scores", 
       y = "Count", 
       fill = "Connectivity Catagory") 


# Histogram for MCDA Z score results 
wsm_Zscore_results2 %>% 
  #filter(connectivity_category == "cascade") %>% 
  ggplot() +
  geom_histogram(aes(wsm_scores, fill = connectivity_category)) + 
  theme_light() + 
  scale_fill_manual(values = connectivity_cat_pal) +
  labs(title = "Zscore WSM Scrores", 
       x = "WSM Scores", 
       y = "Count", 
       fill = "Connectivity Catagory")  
```


C) Table 

```{r}
# Table
wsm_MinMax_results2 %>% 
  st_drop_geometry() %>% 
  group_by(score_letter) %>% 
  summarise(count = n(), 
            avg_scores = mean(wsm_scores, na.rm = TRUE), 
            max_scores = max(wsm_scores, na.rm = TRUE), 
            min_scores = min(wsm_scores, na.rm = TRUE), 
            median_scores = median(wsm_scores, na.rm = TRUE), 
            avg_csi = mean(csi, na.rm = TRUE)) %>% 
  kableExtra::kbl(digits = 3) %>% 
  kableExtra::kable_styling(bootstrap_options = c("striped", "hover", "condensed")) 


```

D) Interactive Map 

For MCDA MinMax results 
```{r}
library(leaflet)
library(leaflet.extras)

# Make it sf
wsm_MinMax_results2_sf <- st_as_sf(wsm_MinMax_results2, crs = 3857)

# Check CRS 
#st_crs(wsm_MinMax_results2_sf) == st_crs(east_as_basin_sf) # Add basin 
wsm_MinMax_results2_sf <- st_transform(wsm_MinMax_results2_sf, 4326) %>% 
  mutate(lon = st_coordinates(.)[,1],
         lat = st_coordinates(.)[,2])


# Make a leaflet
pal <- colorNumeric(
  palette = colorRampPalette(c("#0f31bc", "#77d613", "orange"))(100),
  domain = wsm_MinMax_results2_sf$wsm_scores
)

leaflet(wsm_MinMax_results2_sf) %>%
  addTiles() %>%
  addCircleMarkers(
    lng = ~lon, lat = ~lat,
    color = ~pal(wsm_scores),
    radius = 5,
    fillOpacity = 0.8,
    popup = ~paste0(
      "<b>Dam_ID: ", dam_id, "</b><br>",
      "<b>CSI: </b>", csi, "<br>",
      "<b>Connectivity Category: </b>", connectivity_category_num, "<br>",
      "<b>In a PA: </b>", in_protected_area, "<br>",
      "<b>Capacity: </b>", capacity_mw, " mw"
    )
  ) %>%
  addLegend(pal = pal, values = ~wsm_scores, title = "WSM Scores")
```


```{r}
# Make it sf
wsm_Zscore_results2_sf <- st_as_sf(wsm_Zscore_results2, crs = 3857)

# Check CRS 
#st_crs(wsm_MinMax_results2_sf) == st_crs(east_as_basin_sf) # Add basin 
wsm_Zscore_results2_sf <- st_transform(wsm_Zscore_results2_sf, 4326)  %>% 
  mutate(lon = st_coordinates(.)[,1],
         lat = st_coordinates(.)[,2])


# Map with a leaflet
pal <- colorNumeric(
  palette = colorRampPalette(c("#0f31bc", "#77d613", "orange"))(100),
  domain = wsm_Zscore_results2_sf$wsm_scores
)

leaflet(wsm_Zscore_results2_sf) %>%
  addTiles() %>%
  addCircleMarkers(
    lng = ~lon, lat = ~lat,
    color = ~pal(wsm_scores),
    radius = 5,
    fillOpacity = 0.8,
    popup = ~paste0(
      "<b>Dam_ID: ", dam_id, "</b><br>",
      "<b>CSI: </b>", csi, "<br>",
      "<b>Connectivity Category: </b>", connectivity_category_num, "<br>",
      "<b>In a PA: </b>", in_protected_area, "<br>",
      "<b>Capacity: </b>", capacity_mw, " mw"
    )
  ) %>%
  addLegend(pal = pal, values = ~wsm_scores, title = "WSM Scores")
```


