---
title: "prep_data"
format: html
---

This script preprocesses the dam dataframe output from the Connectivity repository by merging the protected area datasets and applying one-hot encoding to categorical variables, preparing the data for input into the Multi-Criteria Decision Model across all future dam scenarios.

```{r}
# Libraries
library(tidyverse)
library(dplyr)
```

Load data - Dam and protected area data
```{r}
# From the connectivity repository workflow
out_enriched <- readRDS('/capstone/netzerohydro/data/out_enriched/out_enriched_v1.rds') 
out_enriched <- out_enriched[, c(1:17,21, 25:29)]


# From protected area repository workflow
protected_area <- read_csv("/capstone/netzerohydro/data/cleaned_v1/pa_future_dam_global.csv") %>% 
  mutate(dam_id = paste0("FHRED_", dam_id)) %>% # dam_id column needs to match the out_enriched dam_id column 
  select(-c(geometry))
  # st_drop_geometry()

```

## (1) Merge dam and protected area data sets

Add protected area data to each Future dam 
```{r}
# Merge datasets by dam_id 
out_enriched_pa <- left_join(out_enriched, protected_area, by = "dam_id")
```


## (2) Make MCDA criteria numeric 

The Multi-Criteria Decision Model is a mathematical model that requires all criteria must be numeric. 
Therefore, all categorical data must be converted into numbers. Specifically, the connectivity categories must be converted to numbers where the largest number is the most suitable (best) outcome and the smallest number is the least preferred outcome. 

```{r}
# Change connectivity categories to numeric 
out_enriched_pa <- out_enriched_pa %>%
  mutate(connectivity_category_num  = case_when(
    connectivity_category == "cascade_classic" ~ 6, # Most suitable 
    connectivity_category == "cascade1" ~ 5,
    connectivity_category == "cascade2+" ~ 4,
    connectivity_category == "has_downstream_dam" ~ 3,  
    connectivity_category == "has_upstream_dam" ~ 3,
    connectivity_category == "undammed" ~ 1, # Least suitable 
    TRUE ~ 1
    ))
  # If you add any other categorical predictors to the model, you can mutate the columns here 

  # PA was converted into binary in the fatal-flaw workflow. Therefore, code below is not necessary. 
  # %>% mutate(in_pa_num = ifelse(in_protected_area == FALSE, 1, 0)) # 1 = no PA (good for dam placement), 0 = in PA (bad for dam placement)
```

Save the dataframe as it is used in the model and sensitivity testing scripts 
```{r}
# Save merged dataframe in a data folder as it is use in the rest of this repo 
# saveRDS(object = out_enriched_pa, file = here::here("data", "out_enriched_v1.rds"))
```

