---
title: "sensitivity-testing"
format: html
---

### GOAL: See how the dam scores vary with different weights.

```{r}
source(here::here("wsm_model.R"))
```

## Load and Wrangle Data

```{r}
# Load data 
east_as <- readRDS("../../../../../capstone/netzerohydro/data/cleaned_v1/se_as_basin_df.rds")
east_as <- readRDS("../../../../../capstone/netzerohydro/data/cleaned_v1//raf_presentation_df.rds")
east_as_basin_sf <- read_sf("../../../../../capstone/netzerohydro/data/raw/hybas_as_lev02_v1c/hybas_as_lev02_v1c.shp") |>
  janitor::clean_names() %>% 
  filter(hybas_id == 4020015090) %>% 
  st_transform(crs = 4326)

east_as_network <- readRDS("../../../../../capstone/netzerohydro/data/cleaned_v1/nets_with_dams/net_hybas4020015090_Asia_pfaf02.rds")

# Fix dam_id 
future_dams_flagged$dam_id <- as.character(future_dams_flagged$dam_id)

# Join PA with dam data 
east_as_pa <- left_join(east_as, future_dams_flagged, by = 'dam_id')

# Fix the Data 
east_as_pa <- east_as_pa %>%
  mutate(connectivity_category = case_when(
    # (1) Both within 250km
    cascade_level == 1 ~ "cascade1",
    cascade_level == 2 ~ "cascade2",
    cascade_level >= 3 ~ "cascade3+",
    # (2) No dams in either direction OR both further than 200km
    (is.na(min_distance_upstream_km) | min_distance_upstream_km > 70) &
      (is.na(min_distance_downstream_km) | min_distance_downstream_km > 70) ~ "FFR",
    # (3) Upstream dam within 200km, no downstream or downstream > 200km
    min_distance_upstream_km <= 70 &
      (is.na(min_distance_downstream_km) | min_distance_downstream_km > 70) ~ "downstream",
    # (4) Downstream dam within 200km, no upstream or upstream > 200km
    min_distance_downstream_km <= 70 &
      (is.na(min_distance_upstream_km) | min_distance_upstream_km > 70) ~ "upstream",
    TRUE ~ NA
  )) %>% 
  mutate(connectivity_category_num = case_when(
    connectivity_category == "cascade1" ~ 7, 
    connectivity_category == "cascade2" ~ 5, 
    connectivity_category == "cascade3+" ~ 5, 
    connectivity_category == "upstream" ~ 3,
    connectivity_category == "downstream" ~ 3,
    connectivity_category == "FFR" ~ 1,
  )) %>% 
  mutate(in_pa_num = ifelse(in_protected_area == FALSE, 1, 0)) %>% # 1 = No in PA (good)
  #st_drop_geometry() %>% 
  filter(dam_type == "future")
  

```

## Generate Random Weights for the criteria - where each row = 1

Think of it like randomly throwing 2 points on a line from 0 to 1 (for 3 criteria), then measuring the 3 gap sizes — they'll always sum to 1.

```{r}

n_simulations <- 198
n_criteria <- length(criteria)

# EMPTY MATRIX: one row per simulation, one column per criteria 
weight_samples <- matrix(NA, nrow = n_simulations, ncol = n_criteria)
colnames(weight_samples) <- criteria


# FILL MATRIX: with weights - each row = 1 
for (i in 1:n_simulations) {
  w <- diff(c(0, sort(runif(n_criteria - 1)), 1)) # draw (n_criteria - 1) random numbers between 0 and 1  
  weight_samples[i, ] <- w # Find the distance between the (n_criteria - 1) random numbers = random weights 
}

# Make the matrix a list of each criteria combo 
weight_samples_list <- lapply(1:nrow(weight_samples), # Create list from 1 to # of rows in weight_samples 
                               function(i) weight_samples[i, ] )# Function that pulls row by row

```

## Run samples for the dataset over the differing weights

```{r}

criteria <- c('csi', 'connectivity_category_num', 'in_pa_num')

criteria_type <- c(csi = "nonbeneficial", # Lower number = better 
                       connectivity_category_num = "beneficial", # Higher number = better 
                       in_pa_num = "beneficial")

# Run the function over the varying weights 
score_results <- map(.x = weight_samples_list, ~{
  wsm_model(
    dataset = east_as_pa, 
    criteria = criteria, 
    criteria_weights = .x, # Varying
    criteria_type = criteria_type
  )}
    ) %>% bind_cols() # Bind multiple dataframes by column 


score_results
```

```{r}
# Filter for dam #1 
dam_filter <- score_results 

dam_pivot <- dam_filter %>%
  pivot_longer(
    cols = matches("wsm_scores"),
    names_to = "wsm_score",  # was `names =`, which is wrong
    values_to = "value"
  )

dam_pivot_select <- dam_pivot %>% 
  select(c(dam_id...1, wsm_score, value))

dam_wsm_sum <- dam_pivot_select %>%
  group_by(dam_id...1) %>% 
  summarise(
    maxWSM = max(value), 
    minWSM = min(value), 
    avgWSM = mean(value), 
    sdWSM = sd(value)
  ) 

dam_wsm_sum
```

```{r}
criteria <- c('csi', 'connectivity_category_num', 'in_pa_num')

criteria_type <- list(
  csi = "non-beneficial",
  connectivity_category_num = "beneficial",
  in_pa_num = "beneficial"
)

# Run simulations, keeping only wsm_scores per run
score_results <- map(weight_samples_list, \(w) {
  wsm_model(
    dataset = east_as_pa,
    criteria = criteria,
    criteria_weights = w,
    criteria_type = criteria_type
  )$wsm_scores  # Pull scores only
}) %>%
  bind_cols() %>%   # Each col = one simulation
  setNames(paste0("sim_", 1:n_simulations)) # Col names to simulation # 

# Add dam ID for reference
score_results <- bind_cols( # Glues dam col from east_as_pa to score_results horizontally.
  east_as_pa %>% select(dam_id), 
  score_results
)

score_summary <- score_results %>%
  rowwise() %>% # Work within a row 
  mutate(
    mean_score = mean(c_across(starts_with("sim_"))), # Find mean with all columns that start with 'sim_'
    sd_score   = sd(c_across(starts_with("sim_"))), # Find sd with all columns that start with 'sim_'
    cv         = sd_score / mean_score # coefficient of variation — high CV = weight-sensitive
  ) %>% 
    arrange(desc(cv)) %>% 
  select(c(dam_id, mean_score, sd_score, cv))
```

**All dams are showing very sensitive to weights!**

The only dam showing lower senstivity is:

```{r}
score_summary %>% 
  filter(cv <= 0.30)

east_as_pa %>% filter(dam_id == "1969")
```
