---
title: "sensitivity-testing-Zscore"
format: html
---

### GOAL: See how the dam scores vary with different weights.

```{r}
#| output: false 
# Function 
#source(here::here("R", "wsm_model_zscore.R"))

# Libraries 
library(dplyr)
library(tidyverse)
library(sf)
library(ggplot2)
library(plotly)
```

## Load and Wrangle Data

This data is from running the `connectivity repository` workflow with GDW and FRHeD datasets.

```{r}
# Load the out_enriched_pa from the `prep_data.qmd` repo 
out_enriched <- out_enriched_pa 
# out_enriched <- readRDS(here::here("data", "out_enriched_v1.rds")) %>% filter(dam_type == "future") # If saved it as suggested in the `prep_data.qmd`

# Filter for just future dams 
out_enriched_future <- out_enriched %>% 
  filter(dam_type == 'future') %>% 
  st_drop_geometry()
  #%>% filter(threshold_k == '100') # Can filter for specific threshold

# Filter for specific columns 
dataset_filter <- out_enriched_future %>% 
  select(csi, connectivity_category_num, n_protected_areas) %>% 
  st_drop_geometry() %>% 
  mutate(across(everything(), as.numeric))
```

## Generate Random Weights for the criteria - where each row = 1

Think of it like randomly throwing 2 points on a line from 0 to 1 (for 3 criteria), then measuring the 3 gap sizes — they'll always sum to 1.

```{r}
#| output: false 
n_simulations <- 200

criteria <- c('csi', 'connectivity_category_num', 'n_protected_areas') # This is for the function

n_criteria <- length(criteria)

# EMPTY MATRIX: one row per simulation, one column per criteria 
weight_samples <- matrix(NA, nrow = n_simulations, ncol = n_criteria)
colnames(weight_samples) <- criteria


# FILL MATRIX: with weights - each row = 1 
for (i in 1:n_simulations) {
  w <- diff(c(0, sort(runif(n_criteria - 1)), 1)) # draw (n_criteria - 1) random numbers between 0 and 1  
  weight_samples[i, ] <- w # Find the distance between the (n_criteria - 1) random numbers = random weights 
}

# Make the matrix a list of each criteria combo 
weight_samples_list <- lapply(1:nrow(weight_samples), # Create list from 1 to # of rows in weight_samples 
                               function(i) weight_samples[i, ] )# Function that pulls row by row

```

## Extracting only the scores, then analyze rank stability

```{r}
# Run model over simulations 
criteria_type <- list(
  csi = "nonbeneficial",
  connectivity_category_num = "beneficial",
  n_protected_areas = "beneficial"
)

score_results <- map(weight_samples_list, \(w) {
  wsm_model_zscore(
    dataset = dataset_filter,
    criteria = criteria,
    criteria_weights = w,
    criteria_type = criteria_type
  )$wsm_scores
}) %>%
  bind_cols() %>%
  setNames(paste0("sim_", 1:n_simulations))

# Join results to dataframe 
score_results <- bind_cols(
  out_enriched_future %>% select(c(dam_id, threshold_k)),
  score_results
)

# Summary Stats 
score_summary_sf <- score_results %>%
  rowwise() %>% # Work within a row 
  mutate(
    mean_score = mean(c_across(starts_with("sim_")), na.rm = TRUE), # Find mean with all columns that start with 'sim_'
    sd_score   = sd(c_across(starts_with("sim_")), na.rm = TRUE), # Find sd with all columns that start with 'sim_'
    cv         = sd_score / mean_score, # coefficient of variation — high CV = weight-sensitive - DONT USE because sometimes 0/0 = NaN
    max =  max(c_across(starts_with("sim_"))),
    min = min(c_across(starts_with("sim_"))),
    range  = max - min
  ) %>% 
    arrange(desc(cv)) %>% 
  select(c(dam_id, threshold_k, mean_score, sd_score, cv,max, min, range))

# Drop geometeries 
score_summary <- score_summary_sf %>% 
  st_drop_geometry()

```

## Results

### Score Ranges

```{r}
# Look at the ranges 
ggplot(score_summary) + 
  geom_linerange(aes(y = dam_id, xmin = min, xmax = max), color = "grey") + 
  geom_point(aes(x = max, y = dam_id), color = "blue") + 
  geom_point(aes(x = min, y = dam_id), color = "red") + 
  labs(title = "Ranges for all dams") + 
  theme_light()

# Look at dams with high range - These are sensitive to weights  
score_summary %>% 
  filter(range > 0.5) %>% 
  ggplot() +
  geom_linerange(aes(y = dam_id, xmin = min, xmax = max), color = "grey") + 
  geom_point(aes(x = max, y = dam_id), color = "blue") + 
  geom_point(aes(x = min, y = dam_id), color = "red") + 
  labs(x = "WSM Scores", 
       title = "Dams with Ranges over 0.5") + 
  theme_light()

# Look at dams with low range - these are NOT sensitive 
score_summary %>% 
  filter(range < 0.5) %>% 
  ggplot() +
  geom_linerange(aes(y = dam_id, xmin = min, xmax = max), color = "darkgrey") + 
  geom_point(aes(x = max, y = dam_id), color = "blue") + 
  geom_point(aes(x = min, y = dam_id), color = "red") + 
  labs(x = "WSM Scores", 
       title = "Dams with Ranges under 0.5") + 
  theme_light()

```

### Scores mean and sd

Looking at scores mean and range (deviation) for category at each threshold.

```{r}
# For for the 100 threshold
score_summary %>% 
  filter(threshold_k == '100') %>%  
  ggplot(aes(x = mean_score, y = sd_score, label = dam_id)) +
  geom_point() +
  geom_text(size = 2, vjust = -0.5) + 
  theme_light()+ 
  labs(title = "WSM Mean Scores and Standard Deviation")


# For all thesholds 
left_join(score_summary, out_enriched_future, by = "dam_id") %>% 
  ggplot(aes(x = mean_score, y = sd_score, label = dam_id, color = connectivity_category)) +
  geom_point() +
  #geom_text(size = 2, vjust = -0.5) + # Add dam text 
  facet_wrap(~threshold_k.x) +
  scale_color_manual(values = connectivity_cat_pal) + 
  theme_light()+ 
  labs(title = "WSM Mean Scores and Standard Deviation")

```

## Sensitivity Category

Looking at average scoring results per each category of connectivity.

```{r}
# Filter for dams criteria with low sensitivity  
score_low_sensitivity <- left_join(score_summary, out_enriched_future, by = "dam_id") %>% 
  filter(range < 0.5) %>% 
  select(dam_id, threshold_k.x, mean_score, sd_score, range, connectivity_category, connectivity_category_num, csi, n_protected_areas) %>% 
  rename(threshold_k = threshold_k.x)

# Low sensitive Dams 
score_low_sensitivity
dim(score_low_sensitivity)

# Look at results for low sensitivity dams 
score_low_sensitivity_sum <- score_low_sensitivity %>% 
  filter(threshold_k == '100') %>% 
  group_by(connectivity_category) %>% 
  summarise(count = n(), 
            avg_meanWSM = mean(mean_score), 
            med_meanWSM = median(mean_score), 
            avg_rangeWSM = mean(range))

score_low_sensitivity_sum

```

# Quadrant Plot

Looking at scores mean and range (deviation) for each threshold.

```{r}
mean(score_summary$range)
```

```{r}
dam_sensitivity_point <- score_summary %>%
  # filter(threshold_k == '100') %>% 
  mutate(
    priority   = if_else(mean_score > mean(score_summary$mean_score), "High", "Low"), # Avg mean score 
    sensitivity = if_else(range > mean(score_summary$range), "Sensitive", "Robust"), # Avg range score 
    quadrant   = paste(sensitivity, priority, "Priority")
  ) %>%
  ggplot(aes(x = mean_score, y = range, color = quadrant, label = dam_id)) +
  geom_point(size = 2.5) +
  #scale_color_manual(values = c("#4880db", "#74c98e", "#9060D0", "#d31f7f"))+ 
  scale_color_manual(values = c("#07966b", "#A1C566", "#825898", "#FD96B9"))+ 
  #geom_text(size = 2, vjust = -0.6, alpha = 0.6) +
  geom_vline(xintercept = mean(score_summary$mean_score), linetype = "dashed", color = "grey50") +
  geom_hline(yintercept = mean(score_summary$range), linetype = "dashed", color = "grey50") +
  facet_wrap(~threshold_k) +
  labs(x = "Mean WSM Score", y = "Score Range (max − min)",
       title = "Dam Priority vs. Weight Sensitivity",
       color = NULL) +
  theme_light() + 
  theme(panel.spacing = unit(5, "mm"))

dam_sensitivity_point

```

The large Robust Low Priority (light green) cluster means the majority of dams score consistently low regardless of weights, should not be prioritizes.

The small Robust High Priority (dark green) cluster is the most defensible group with high scores that hold up under any weighting. Dams that score high and stays high no matter how you weight the criteria. These are confidently good candidates. Easy to defend.

The massive Sensitive High Priority (purple) cloud is the biggest issue. These are high-scoring dams that are highly sensitive. These dams score well on average, but their range is near 0.5–1.0, meaning weights dramatically change their scores. You can't confidently prioritize them without justifying your weight choices.

The Sensitive Low Priority (pink) cluster is similarly unstable. These scores low on average, but could score high under certain weight combinations.
